---
title: "06_HSA_hotspots"
author: "Clara Peiret-Garcia"
date: "2023-08-29"
output: html_document
---

# Introduction
Detect hot and cold-spots based on the accessibility indexes.

```{r}
# Libraries
library(sf)
library(sfdep)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(ggnewscale)
library(ggspatial)
library(ggpubr)

```



```{r}
# Data
scores_oa <- readRDS('./outputs/scores_oa.RData')
```

```{r}
# Generate a function that performs the hot-spot analysis
hs_analysis <- function(data, geometry_column, attribute_column) {
  # Step 1: Create nb and wt columns
  data <- data %>%
    mutate(
      nb = st_contiguity(!!rlang::sym(geometry_column)),
      wt = st_weights(nb, allow_zero = TRUE)
    )
  
  # Step 2: Calculate A_lag
  data <- data %>%
    mutate(
      A_lag = st_lag(!!rlang::sym(attribute_column), nb, wt, allow_zero = TRUE)
    )
  
  # Step 3: Calculate local G statistics
  data <- data %>%
    mutate(
      gi = local_g_perm(!!rlang::sym(attribute_column), nb, wt, nsim = 500)
    ) %>%
    unnest(gi)
  
  # Step 4: Classify the outputs
  data <- data %>%
    dplyr::select(OA21CD, gi, p_folded_sim) %>%
    mutate(
      classification = case_when(
        gi > 0 & p_folded_sim <= 0.01 ~ 'Hot spot - 99% Confidence',
        gi > 0 & p_folded_sim <= 0.05 ~ 'Hot spot - 95% Confidence',
        gi > 0 & p_folded_sim <= 0.1 ~ 'Hot spot - 90% Confidence',
        gi < 0 & p_folded_sim <= 0.01 ~ 'Cold spot - 99% Confidence',
        gi < 0 & p_folded_sim <= 0.05 ~ 'Cold spot - 95% Confidence',
        gi < 0 & p_folded_sim <= 0.1 ~ 'Cold spot - 90% Confidence',
        TRUE ~ 'Not significant'
      ),
      classification = factor(
        classification,
        levels = c('Hot spot - 99% Confidence',
                   'Hot spot - 95% Confidence',
                   'Hot spot - 90% Confidence',
                   'Not significant',
                   'Cold spot - 90% Confidence',
                   'Cold spot - 95% Confidence',
                   'Cold spot - 99% Confidence')
      )
    )
  
  return(data)
}
```

# Hot-spot analysis for each scenario

```{r}
# Create an empty list to store the results
hs_results <- list()

# Iterate through each data frame in scores_oa
for (i in 1:length(scores_oa)) {
  result <- hs_analysis(data = scores_oa[[i]], 
                        geometry_column = 'geometry', 
                        attribute_column = 'A')
  # Assign names to the hs_results elements
  result_name <- names(scores_oa)[i]
  hs_results[[result_name]] <- result
}
```

# Visualise results

```{r}
hs_maps <- list()
  for (i in 1:length(hs_results)) {
    hs_maps[[i]] <- ggplot(data = hs_results[[i]], aes(fill=classification), lwd = NA, col = NA) +
  geom_sf(lwd = 0, col = NA) +
  theme_minimal() +
  #theme(legend.position = 'none') +
  scale_fill_brewer(type = 'div', palette = "Spectral", 'Classification') +
  annotation_scale(style = "ticks", height = unit(0.15, "cm")) +
  annotation_north_arrow(style = north_arrow_minimal, location = "tl", height = unit(0.75, "cm")) +
  ggtitle(names(hs_results[i])) 
  }

#grid.arrange(grobs = hs_maps, ncol = 2)

pdf("./figures_paper/HSA_maps.pdf", paper = "USr", height = 8.5, width = 11)
ggarrange(plotlist = hs_maps, common.legend = TRUE, legend = "bottom")
dev.off()
```


```{r}
# Create stack plots
# Transform crs
transformed <- hs_results$Baseline

rotate_data <- function(data, x_add = 0, y_add = 0) {

  shear_matrix <- function(){ matrix(c(2, 1.2, 0, 1), 2, 2) }

  rotate_matrix <- function(x){
    matrix(c(cos(x), sin(x), -sin(x), cos(x)), 2,2)
  }
  data %>%
    dplyr::mutate(
      geometry = .$geometry * shear_matrix() * rotate_matrix(pi/20) + c(x_add, y_add)
    )
}

# Plots
x = 70.1
color = 'gray40'
custom_labels <- c('Very hot', 'Hot', 'Somewhat hot', 'Non-significant', 'Somewhat cold', 'Cold', 'Very cold')


temp1 <- ggplot() +

  # Students
  geom_sf(data = hs_results$Students %>% rotate_data(),  aes(fill=classification), lwd = 0) +
  scale_fill_brewer(type = 'div', palette = "Spectral", 'Classification') +
  annotate("text", label='Students', x=x, y= 44.5, hjust = 0, color=color)


temp2 <- temp1 +
  new_scale_fill() +
  new_scale_colour() +

  # Active life

  geom_sf(data = hs_results$`Active life` %>% rotate_data(y_add = 0.15),  aes(fill=classification), lwd = 0) +
  annotate("text", label='Active life', x=x, y= 44.5 + 0.15 , hjust = 0, color=color) +

  # Work-life balance
  geom_sf(data = hs_results$`Work-life balance` %>% rotate_data(y_add = 0.3),  aes(fill=classification), lwd = 0) +
  annotate("text", label='Work-life balance', x=x, y= 44.5 + 0.3 , hjust = 0, color=color) +

  # Culture fans
  geom_sf(data = hs_results$`Culture fans` %>% rotate_data(y_add = 0.45),  aes(fill=classification), lwd = 0) +
  annotate("text", label='Culture fans', x=x, y= 44.5 + 0.45 , hjust = 0, color=color) +

  # Housekeepers
  geom_sf(data = hs_results$Housekeepers %>% rotate_data(y_add = 0.6),  aes(fill=classification), lwd = 0) +
  annotate("text", label='Housekeepers', x=x, y= 44.5 + 0.6 , hjust = 0, color=color) +

  scale_fill_brewer(type = 'div', palette = "Spectral", 'Classification') +
  theme_void()

temp2

```


# Hot-spot persistence
We now want to analyse if there are areas that are consistently hot/cold-spots. We also want to see which areas are not considered hot/cold-spots under the baseline scenario, but show up as hot/cold-spots for the alternative scenarios.

For the hot-spot persistence, we only consider the alternative scenarios, and not the baseline. 

```{r}
# Hot-spot persistence
# Remove baseline scenario
hs_results_b <- hs_results[-1]

```

```{r}
# Combine all data frames into a single data frame
combined_df <- bind_rows(hs_results_b, .id = "Scenario")

# Custom function to calculate the mode
calculate_mode <- function(x) {
  uniq_x <- unique(x)
  freq_x <- tabulate(match(x, uniq_x))
  uniq_x[which.max(freq_x)]
}

# Group by observation and calculate mode and its frequency using the custom function
result <- combined_df %>%
  group_by(OA21CD) %>%
  summarise(
    mode_classification = calculate_mode(classification),
    mode_frequency = max(tabulate(match(classification, calculate_mode(classification))))
  )

head(result)

```

```{r}
# Convert classification belonging and frequency into a single variable

result <- result %>% 
  mutate(
    class_numeric = case_when(
      mode_classification %in% c("Hot spot - 99% Confidence") ~ 3,
      mode_classification %in% c("Hot spot - 95% Confidence") ~ 2,
      mode_classification %in% c("Hot spot - 90% Confidence") ~ 1,
      mode_classification %in% c("Cold spot - 90% Confidence") ~ -1,
      mode_classification %in% c("Cold spot - 95% Confidence") ~ -2,
      mode_classification %in% c("Cold spot - 99% Confidence") ~ -3,
      TRUE ~ 0
    )
  )

# Now, multiply persistence value by the times that value occurs for each OA
result$persistence_value <- result$class_numeric * result$mode_frequency

# Normalise values from -1 to 1.
min_value <- min(result$persistence_value)
max_value <- max(result$persistence_value)

result$norm_persistence_value <- 2 * (result$persistence_value - min_value) / (max_value - min_value) - 1


```


```{r}

# Plot persistence
p_persistence <- ggplot() +
  geom_sf(data = result, aes(fill = persistence_value), lwd = NA) +
  scale_fill_distiller(palette = "Spectral", "Persistence") +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  annotation_scale(style = "ticks", height = unit(0.15, "cm"), text_cex = 0.75) +
  annotation_north_arrow(style = north_arrow_minimal, location = "tl", height = unit(0.75, "cm")) +
  ggtitle("Hot-spot persistence across scenarios")
  #theme(legend.text = element_text(size = 16),
        #legend.title = element_text(size = 18))
   
```

```{r}
# Peristence compared to baseline scenario

ggarrange(p1, p_persistence)

```


# Scenario comparison

We want to know how much each individual scenario differs from the baseline. To do so, we transform the categorical classification of hotspots into a numerical classification, and calculate the numerical differences between both.

```{r}

# Create a new variable that transforms categorical values into numerical.
hs_results <- lapply(hs_results, function(x){
  x %>% 
  mutate(classification_numeric = case_when(
   classification %in% c("Hot spot - 99% Confidence") ~ 3,
   classification %in% c("Hot spot - 95% Confidence") ~ 2,
   classification %in% c("Hot spot - 90% Confidence") ~ 1,
   classification %in% c("Cold spot - 90% Confidence") ~ -1,
   classification %in% c("Cold spot - 95% Confidence") ~ -2,
   classification %in% c("Cold spot - 99% Confidence") ~ -3,
   TRUE ~ 0
  )
  )
})

# Calculate difference between each scenario and the baseline
hs_results <- lapply(hs_results, function(x){
  x %>%   mutate(dif_baseline = hs_results$Baseline$classification_numeric - x$classification_numeric) 
})

```

## Explore results

```{r}
baseline_dif_maps <- list()
  for (i in 2:length(hs_results)) {
    baseline_dif_maps[[i]] <- ggplot(data = hs_results[[i]], aes(fill=dif_baseline), lwd = NA, col = "white") +
  geom_sf(lwd = 0, col = 0) +
  theme_minimal() +
  #theme(legend.position = 'none') +
  scale_fill_distiller(palette = "PiYG", direction = -1,"Difference with baseline") +
  annotation_scale(style = "ticks", height = unit(0.15, "cm")) +
  annotation_north_arrow(style = north_arrow_minimal, location = "tl", height = unit(0.75, "cm")) +
  ggtitle(names(hs_results[i])) 
  }

baseline_dif_maps[[1]] <- NULL

#grid.arrange(grobs = baseline_dif_maps, ncol = 2)
#pdf("./figures_paper/HSA_baseline_dif.pdf", paper = "USr", height = 8.5, width = 11)
ggarrange(plotlist = baseline_dif_maps, common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 2)
#dev.off()
```

```{r}
# Do the same but for accessibility scores
# Calculate difference between each scenario and the baseline
scores_oa <- lapply(scores_oa, function(x){
  x %>%   mutate(dif_baseline = scores_oa$Baseline$A - x$A) 
})

```

```{r}
# Map
access_baseline_dif_maps <- list()
  for (i in 2:length(scores_oa)) {
    access_baseline_dif_maps[[i]] <- ggplot(data = scores_oa[[i]], aes(fill=dif_baseline), , lwd = NA, col = NA) +
  geom_sf(lwd = 0, col = 0) +
  theme_minimal() +
  #theme(legend.position = 'none') +
  scale_fill_viridis_c(option = "G", direction = -1,"Difference with baseline") +
  annotation_scale(style = "ticks", height = unit(0.15, "cm")) +
  annotation_north_arrow(style = north_arrow_minimal, location = "tl", height = unit(0.75, "cm")) +
  ggtitle(names(hs_results[i])) 
  }

access_baseline_dif_maps[[1]] <- NULL

ggarrange(plotlist = access_baseline_dif_maps, common.legend = TRUE, legend = "bottom")

```














